% LaTeX Curriculum Vitae Template
%
% Copyright (C) 2004-2009 Jason Blevins <jrblevin@sdf.lonestar.org>
% http://jblevins.org/projects/cv-template/
%
% You may use use this document as a template to create your own CV
% and you may redistribute the source code freely. No attribution is
% required in any resulting documents. I do ask that you please leave
% this notice and the above URL in the source code if you choose to
% redistribute this file.

\documentclass[letterpaper]{article}

\usepackage{hyperref}
\usepackage{geometry}

% Comment the following lines to use the default Computer Modern font
% instead of the Palatino font provided by the mathpazo package.
% Remove the 'osf' bit if you don't like the old style figures.

% Set your name here
\def\name{Kris Sankaran}

% Replace this with a link to your CV if you like, or set it empty
% (as in \def\footerlink{}) to remove the link in the footer:
\def\footerlink{}
\usepackage[dvipsnames]{xcolor}

\newcommand\myshade{85}
\colorlet{mylinkcolor}{violet}
\colorlet{mycitecolor}{YellowOrange}
\colorlet{myurlcolor}{Aquamarine}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  linkcolor = mylinkcolor!\myshade!black,
  citecolor = mycitecolor!\myshade!black,
  urlcolor = myurlcolor!\myshade!black,
  pdfauthor = {\name},
  pdfkeywords = {economics, statistics, mathematics},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}

\geometry{
  body={6.5in, 8.5in},
  left=0.7in,
  top=0.7in,
  bottom=0.9in
}

% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\thispagestyle{empty}

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

% Make lists without bullets
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
  }
}{
  \end{list}
}

\begin{document}

% Place name at left
{\huge \name}

% Alternatively, print name centered and bold:
%\centerline{\huge \bf \name}

\vspace{0.25in}

\begin{minipage}{0.45\linewidth}
  Qu\'ebec Artificial Intelligence Institute \\
  6666 Rue Saint Urbain \\
  Montr\'eal, QC \\
  H2S 3H1
\end{minipage}
\begin{minipage}{0.45\linewidth}
  \begin{tabular}{ll}
    Email: & \href{mailto:sankaran.kris@gmail.com}{\tt sankaran.kris@gmail.com} \\
    Homepage: & \href{http://krisrs1128.github.io/personal-site/}{\tt http://krisrs1128.github.io/personal-site/} \\
    Code: & \href{https://github.com/krisrs1128}{\tt https://github.com/krisrs1128}
  \end{tabular}
\end{minipage}

\section*{Education}
Stanford University
\begin{itemize}
  \item PhD Statistics, 2013 - 2018
    \begin{itemize}
    \item Thesis: \textit{Discovery and Visualization of Latent Structure with Applications to the Microbiome}
    \end{itemize}
  \item MS Statistics, 2012 - 2013
  \item BS Mathematics, 2009 - 2013
\end{itemize}

\section*{Professional Experience}
\begin{itemize}
\item \textit{Postdoctoral Fellow}\\
August 2018 - Present \\
D\'epartement d'Informatique et de Recherche Op\'erationnelle \\
Universit\'e de Montr\'eal

\item \textit{Senior Machine Learning Researcher} \\
  April 2018 - August 2018 \\
  Obsidian Security
\item \textit{Data Science for Social Good Fellowship} \\
  Summer 2016 \\
  Center for Data Science and Public Policy, University of Chicago
\item \textit{Quantitative Research Intern} \\
  Summer 2012 \\
  The Climate Corporation
\end{itemize}

\section*{Teaching Experience}
\begin{itemize}
\item \textit{IFT6758 (Introduction to Data Science) co-instructor} \\
Fall 2019 \\
D\'epartement d'Informatique et de Recherche Op\'erationnelle \\
Universit\'e de Montr\'eal
\item \textit{Data Science Nigeria AI Bootcamp instructor} \\
Session on ``Humanitarian AI.'' \\
Lagos, Nigeria. \\
November 2019.
\item \textit{Deep Learning Indaba instructor} \\
Session on ``Memory and Attention.'' \\
Nairobi, Kenya. \\
August 2019
\item \textit{First Nepal Winter School in AI instructor}. \\
Sessions on ``Deep Learning Foundations,'' ``Advanced Deep Learning,'' and ``Humanitarian AI.'' \\
Kathmandu, Nepal. \\
December 2018.
\item \textit{Stats 390 (Free Statistical Consulting Workshop) instructor} \\
  Summer 2017 \\
  Stanford University Department of Statistics
\item \textit{Strategies and Techniques for Analyzing Microbial Population Structures TA} \\
  August 2017 \\
  Marine Biological Laboratory, Woods Hole
\end{itemize}

\section*{Areas of Expertise}
\begin{itemize}
  \item Data, augmentation, and inference for the public good
  \item Statistical and deep learning
  \item Heterogeneous and high-dimensional data
  \item Data visualization
\end{itemize}
\section*{Peer-Reviewed Publications}
\begin{itemize}
  \item Methods
\begin{itemize}
\item Sankaran, Kris, and Susan Holmes. Multitable Methods for Microbiome Data Integration. Frontiers in Genetics 10 (2019): 627.
\item Kris Sankaran. Remembrances of States Past. 2nd Workshop on
Visualization for AI Explainability at IEEE VIS 2019. 
\item  Huang, Chin-Wei, Kris Sankaran, Eeshan Dhekane, Alexandre Lacoste, Aaron Courville. Hierarchical Importance Weighted Autoencoders.
Proceedings of the 36th International Conference on Machine Learning, PMLR 97:2869-2878, 2019.
 \item Sankaran, Kris, and Susan P. Holmes. Latent Variable Modeling for the
   Microbiome. Biostatistics, June 3, 2018.
   https://doi.org/10.1093/biostatistics/kxy018
 \item Sankaran, Kris, and Susan P. Holmes. Interactive Visualization of
   Hierarchically Structured Data. Journal of Computational and Graphical
   Statistics, October 18, 2017. \\
   https://doi.org/10.1080/10618600.2017.1392866
\item Sankaran, Kris, and Susan P. Holmes. structSSI: Simultaneous and Selective
  Inference for Grouped or Hierarchically Structured Data. Journal of
  Statistical Software 59, no. 13 (2014).
\end{itemize}
\item Applications
  \begin{itemize}
  \item Gonzalez Villareal, Jessenia, Debjani Bhowmick, and Kris Sankaran. Segmentation of water bodies in Peru using knowledge-transfer based convolutional neural networks. Workshop on ML for the Developing World at NeurIPS 2019.
  \item Goytom, Israel and Kris Sankaran. Forecasting Maxima in Climate Time Series. 9th International Workshop on Climate Informatics. October 2019.
  \item Yuan, Tianle, Hua Song, David Hall, Victor Schmidt, Kris Sankaran and Yoshua Bengio. Artificial intelligence based cloud distributor (AI-CD): probing clouds with generative adversarial networks. 9th International Workshop on Climate Informatics. October 2019.
\item Janik, Adrianna and Kris Sankaran. Discovering Concepts in Learned Representations using Statistical Inference and Interactive Visualization. Workshop on Explainable AI/ML (XAI) for Accountability, Fairness, and Transparency at KDD 2019. https://github.com/adrijanik/adrijanik.github.io/.
  \item Razzak, Muhammed and Kris Sankaran. Interactive Segmentation for Disaster Relief Mapping [Oral presentation]. AI for Good Workshop at IJCAI 2019.
  \item Janik, Adrianna, Kris Sankaran, and Anthony Ortiz. Interpreting Black-Box Semantic Segmentation Models in Remote Sensing Applications. In \textit{Machine Learning Methods in Visualisation for Big Data} edited by Archambault, Daniel and Nabney, Ian and Peltonen, Jaakko (2019). DOI 10.2312/mlvis.20191158.
  \item Benabbou, Loubna, Zouheir Malki, Kris Sankaran, and Hicham Bouzeckri. Machine Learning based Predictive Maintenance for Renewable Energy: The Case of Power Plants in Morocco. Climate Change AI workshop at ICML 2019.
  \item Schmidt, Victor, Alexandra Luccioni, Surya Mukkavilli, Narmada Balasooriya, Kris Sankaran, Jennifer Chayes, and Yoshua Bengio. (2019). AI for Social Good Workshop at ICLR 2019. Visualizing the consequences of climate change using cycle-consistent adversarial networks. arXiv preprint arXiv:1905.03709.
  \item Min, Yan, Xiaoguang Ma, Kris Sankaran, Yuan Ru, Lijin Chen, Mike Baiocchi, and Shankuan Zhu (2019). Sex-specific association between gut microbiome and fat distribution. Nature communications, 10(1), 2408.
  \item Bastidas, Alexei, Matthew Beale, Yoshua Bengio, Anna Bethke, Pablo Fonseca, Jason Jo, Dale Kunce, Sean McPherson, Vincent Michalski, Anthony Ortiz, Kris Sankaran, and Hanlin Tang. Foundational mapping of Uganda to assist American Red Cross disaster response to floods and pandemics. AI for Social Good Workshop at NeurIPS 2019.
  \item Fukuyama, Julia, Laurie Rumker, Kris Sankaran, Pratheepa Jeganathan, Les
    Dethlefsen, David A. Relman, and Susan P. Holmes. Multidomain analyses of a
    longitudinal human microbiome intestinal cleanout perturbation experiment.
    PLOS Computational Biology 13, no. 8 (2017): e1005706.
  \item Callahan, Ben J., Kris Sankaran, Julia A. Fukuyama, Paul J. McMurdie, and
    Susan P. Holmes. Bioconductor workflow for microbiome data analysis: from
    raw reads to community analyses. F1000Research 5 (2016).
  \item Rhee, Soo-Yon, Kris Sankaran, Vici Varghese, Mark Winters, Christopher B.
    Hurt, Joseph J. Eron, Neil Parkin, Susan P. Holmes, Mark Holodniy, and Robert
    W. Shafer. HIV-1 Protease, Reverse Transcriptase, and Integrase Variation.
    Journal of Virology (2016): JVI-00495.
  \item Sankaran, Siva, and Kris Sankaran. Improving Online Course Performance
    Through Customization: An Empirical Study Using Business Analytics.
    International Journal of Business Analytics (IJBAN) 3, no. 4 (2016): 1-20.
  \end{itemize}
\end{itemize}

\section*{Unpublished Reports}
\begin{itemize}
\item Sankaran, Kris. ``Societal Impacts'' in 
\textit{Tackling Climate Change with Machine Learning.} Collection with Rolnick, D., Donti, P. L., Kaack, L. H., Kochanski, K., Lacoste, A., Sankaran, K., ... Luccioni, A. (2019). arXiv preprint arXiv:1906.05433. arXiv preprint arXiv:1906.05433.
\item Casavant, Ellen, Les Dethlefsen, Kris Sankaran, Dan Sprockett, Susan P. Holmes, David Relman, and Joshua Elias. (2019). Strategies for understanding dynamic, personalized profiles of host-derived proteins and microbes from human stool. bioRxiv, 551143.
 \item Sankaran, Kris, and Susan P. Holmes. Inference of Dynamic Regimes in
   the Microbiome. arXiv preprint arXiv:1712.00067 (2017).
  \item Sankaran, Kris, Diego Garcia-Olano, Mobin Javed, Maria Fernanda
    Alcala-Durand, Adolfo De Unánue, Paul van der Boor, Eric Potash, Roberto
    S\'anchez Avalos, Luis Inaki Alberro Encinas, and Rayid Ghani. Applying
    Machine Learning Methods to Enhance the Distribution of Social Services in
    Mexico. arXiv preprint arXiv:1709.05551 (2017).
  \item Sankaran, Kris, Suzanne Tamang, and Ami Bhatt. Opioid Atlas: Mapping
    Access to Pain Medication. arXiv preprint arXiv:1612.00497 (2016).
\end{itemize}

\section*{Ongoing Projects}
Manuscripts available upon request.
\begin{itemize}
\item HighRes-net: Multiframe Super-Resolution by Recursive Fusion.  
\href{https://github.com/ElementAI/HighRes-net}{Code}.
\end{itemize}

\section*{Supervision}
Interns supervised (Mila): Anthony Ortiz, Md. Rifat Arefin, Pablo Fonseca, Muhammed Razzak, Israel Goytom, Adrianna Janik, Yasser Salah Eddine Bouchareb, Sara Ebrahim, Mustafa Alghali, Jessenia Gonzalez Villareal, Wisdom d'Almeida, Shimaa Baraka, Debjani Bhowmick, and Joseph Baafi. \\
Interns co-supervised (Mila): The \href{https://sites.google.com/view/biaslyai/home}{biasly AI} team.

\section*{Funding}
AI for Earth Microsoft Azure Compute Grant [\$18K] for
``Glacier Monitoring for Adaptation to Climate Change.'' October 2019.

\section*{Service}

\begin{itemize}
\item Reviewer for \textit{Biostatistics}, \textit{The Journal of the American Statistical Association}, and workshops on \textit{AI for Social Good} [NeurIPS 2018, ICLR 2019, ICML 2019], \textit{AI for Good} [IJCAI 2019], \textit{ML for Development} [NeurIPS 2019], \textit{Montreal AI Symposium} [2019], \textit{AI for Social Impact} [AAAI 2020], and metareviewer for \textit{Climate Change AI} [ICML 2019, NeurIPS 2019],\\
  2016 - Present
\item  On steering committee for Climate Change AI. \href{https://climatechange.ai}{climatechange.ai}
\item Workshop organizer for \textit{Computer Vision for Global Challenges} [CVPR 2019], \textit{AI for Social Good} [ICML 2019, ICLR 2019]
\item Red Judge for IBM Watson AI XPRIZE \\
May - October 2019.
\item DataKind SF Chapter Core Volunteer \\
  2016 - 2017
\item Stanford Statistics for Social Good \\
  2013 - 2017
\item Stats 390 (Free Statistical Consulting Workshop) participant \\
  13 quarters, 2012 - 2017
\end{itemize}

\section*{Invited Talks}
\begin{itemize}
  \item University of Chicago, Department of Statistics Colloquium (01/2018)
\end{itemize}

\section*{Recognition}
\begin{itemize}
\item Best reviewer award. AI for Social Good Workshop, NeurIPS 2018.
\item Jerome H. Friedman Applied Statistics Dissertation Award 2018
\item Ric Weiland Graduate Fellowship, 2016 - 2018
\item \href{https://undergrad.stanford.edu/programs/pwr/publications-prizes-and-awards/boothe-prize-essays}{Boothe Prize} for Excellence in Writing (Top 6 freshman papers), 2010
\end{itemize}

\begin{center}
  \begin{footnotesize}
    Last updated: \today \\
    \href{\footerlink}{\texttt{\footerlink}}
  \end{footnotesize}
\end{center}

\end{document}
